---
title: "Homework 02"
format: 
  html: 
    embed-resources: true
---


## TidyTuesday Section (optional)

::: {.callout-important title="Instructions"}

You can count work on this week's TidyTuesday toward the exceptional work required for an A in the Homework component.

:::


Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

```{r}
library(tidyverse)
edible_plants <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-02-03/edible_plants.csv')

```
```{r}
edible_plants_clean <- edible_plants %>%
  mutate(sunlight = recode(sunlight, 
         'Full sun/partial shade/ full shade' = 'Full sun/partial shade/full shade',
         'full sun/partial shade/ full shade' = 'Full sun/partial shade/full shade',
         'partial shade' = 'Partial shade')) %>%
  mutate(diff = preferred_ph_upper - preferred_ph_lower)

```

```{r fig.alt= "How Particular is Your Food? Edible plant species vary by desired pH range and sunlight level. Most plants prefer full sun and have a range of about 1 pH level of tolerance."}
ggplot(edible_plants_clean, aes(x = reorder(taxonomic_name, diff), color = sunlight)) +
  geom_linerange(aes(ymin = preferred_ph_lower, ymax = preferred_ph_upper), linewidth = 1) +
  scale_color_viridis_d(name = "Sunlight Level") +
  labs(x = " ", y = "Preferred pH", title = "How Particular Is Your Food?", subtitle = "Edible Plant Species by Preferred Soil pH and Sunlight Level", caption = "via TidyTuesday, by Sam Kenney")+
  theme_classic() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

```
Most edible plants enjoy pH between 5 and 8. Some have larger ranges of up to 3 pH levels, but most tolerate around 1 level of pH. Very few plants prefer to have partial shade, with most desiring full sun or a mix, but there does not appear to be a clear relationship between pH preference and sunlight preference. 
