---
title: "Homework 03"
format: 
  html: 
    embed-resources: true
---


## TidyTuesday Section (optional)

::: {.callout-important title="Instructions"}

You can count work on this week's TidyTuesday toward the exceptional work required for an A in the Homework component.

:::


Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

```{r}
library(tidyverse)
olympics_schedule <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-02-10/schedule.csv')

```

```{r}
olympics_schedule %>%
  filter(is_medal_event == TRUE & is_training == TRUE)
```

```{r}
olympics_schedule_sums <- olympics_schedule %>%
  group_by(discipline_name, is_medal_event, date, is_training) %>%
  summarize(count = n(), .groups = "drop") %>%
  mutate(status = case_when(
    is_training == "TRUE" ~ "training",
    is_medal_event == "TRUE" ~ "medal event", 
    .default = "standard event"
  ))

head(olympics_schedule_sums)
```

```{r}
ggplot(olympics_schedule_sums, aes(x = date, y = discipline_name, color = status, size = count)) +
  scale_color_viridis_d(name = "Event Type") +
  geom_point(alpha = 0.75) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "What's the Best Day to Go to the Olympics?", subtitle = "Event type by count, discipline, and day of the week.") 
  

```
